Stage 2 Yield Prediction

### 2016-02-10 Rmd/empirical-yields.Rmd
Updated on DSSAT and empirical modeling side: 1. Created DSSAT-yields.Rmd, for running DSSAT for both maize and soybean. 2. Results of these go into yield-predictors and empirical-yields to create modeled crop yields (using a GAM). 3. zmd-gsod-stations.Rmd creates the station points used in the analysis.

p_root <- lmisc::set_base_path()
p_clim <- fp(p_root, "external/input_devel/climate")
p_yield <- fp(p_root, "external/input_devel/yield")

@@

# Fit model
## Exploratory plots
```{r, eval = FALSE}
plot(prov)
regdt[CROP == "maize", points(x, y, pch = 16, cex = 0.5)]
regdt[CROP == "maize" & YLDS > 8000, 
      points(x, y, pch = 16, cex = 0.5, col = "red")]
regdt[CROP == "maize" & YLDS > 8000, 
      points(x, y, pch = 16, cex = 0.5, col = "red")]
regdt[CROP == "soybean" & YLDS > 1700, 
      points(x, y, pch = 16, cex = 0.5, col = "blue")]

# Prune out yields > 8000 maize > 1700 soybean in subsistence case
mzfilt <- regdt[CROP == "soybean" & YLDS > 1700, unique(.SD), 
                .SDcols = c("x", "y", "SOL", "WTH", "Zone")]
soyfilt <- regdt[CROP == "maize" & YLDS > 8000, unique(.SD), 
                 .SDcols = c("x", "y", "SOL", "WTH", "Zone")]
regnew <- regdt[!x %in% soyfilt$x | !y %in% soyfilt$y]
# regnew[CROP == "maize", points(x, y, pch = 16, cex = 0.5, col = "green")]

@@

# summary(lm(YLDS ~ poly(MAP, 2), data = regnew[CROP == "maize", ]))
# summary(lm(YLDR ~ poly(MSP, 2), data = regnew[CROP == "maize", ]))
regnew[CROP == "maize", plot(MAP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MAP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MAP, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MSP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MSP, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(FP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(MSP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(MSP, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(sloc, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(sloc, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(slcl, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(slcl, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(slsi, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(slsi, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(sbdm, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(sbdm, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(slhw, YLDS, pch = 16, cex = 0.5)]  # pH looks pred
regnew[CROP == "maize", plot(slhw, YLDR, pch = 16, cex = 0.5)]  # pH looks pred
regnew[CROP == "maize", plot(GDD, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(KDD, YLDS, pch = 16, cex = 0.5)]
# regnew[CROP == "maize", plot(log(KDD), YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(NKJ, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(NKF, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(ATMX, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(STMX, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(STMX, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SSRT, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SSRM, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(ASRT, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(ASRM, YLDS, pch = 16, cex = 0.5)]

@@

summary(lm(YLDS ~ poly(sbdm, 2) + poly(slhw, 2), 
           data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(slhw, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(sbdm, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(KDD, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDR ~ poly(STMX, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(STMX, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ KDD, data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(GDD, 2), data = regnew[CROP == "maize", ]))

# p <- "sbdm"; d <- "YLDS"
# regnew[CROP == "maize", {
#   lm1 <- lm(get(d) ~ get(p))
#   lm2 <- lm(get(d) ~ poly(get(p), 2))
#   plot(get(p), get(d), pch = 16, cex = 0.1)
#   abline(lm1)
#   dum <- data.frame(floor(min(get(p))):ceiling(max(get(p))))
#   colnames(dum) <- p
#   lines(dum[, 1], predict(lm2, newdata = dum), col = "red")
#   summary(lm1)
#   # summary(lm2)
# }]

reg_plotter <- function(dep, pred) {
  lm1 <- lm(dep ~ pred)
  lm2 <- lm(dep ~ poly(pred, 2))
  plot(pred, dep, pch = 16, cex = 0.1)
  abline(lm1)
  prng <- c(floor(min(pred)), ceiling(max(pred)))
  dum <- data.frame(seq(prng[1], prng[2], by = diff(prng) / 100))
  # dum <- data.frame(floor(min(pred)):ceiling(max(pred)))
  colnames(dum) <- "pred"
  preds <- predict(lm2, newdata = dum)
  lines(dum[, 1], preds, col = "red")
  mtext(round(summary(lm1)$adj.r.squared, 4), side = 3, adj = 0)
  mtext(round(summary(lm2)$adj.r.squared, 4), side = 3, adj = 0.5)
}
# regnew[CROP == "maize", reg_plotter(YLDR, GDD)]  # 
regnew[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.045
regnew[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.039
regnew[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.012
# regnew[CROP == "maize", reg_plotter(YLDS, JP - mean(JP))]  # P 0.012
regnew[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.005
regnew[CROP == "maize", reg_plotter(YLDS, OP)]  # P 0.003

regnew[CROP == "maize", reg_plotter(YLDS, slhw)]  # P 0.17
regnew[CROP == "maize", reg_plotter(YLDS, sloc)]  # P 0.266
# regnew[CROP == "maize", reg_plotter(YLDR, sloc)]  # P 
regnew[CROP == "maize", reg_plotter(YLDS, slcl)]  # P 0.185
regnew[CROP == "maize", reg_plotter(YLDS, sbdm)]  # 0.181
# mtax pairings
regnew[CROP == "maize", reg_plotter(YLDS, STMX)]  # P 0.07
regnew[CROP == "maize", reg_plotter(YLDS, ATMX)]  # P 0.056 
# NK
regnew[CROP == "maize", reg_plotter(YLDS, NKJ)]  # L 0.054
regnew[CROP == "maize", reg_plotter(YLDS, NKF)]  # P 0.027 
# GDD/KDD
regnew[CROP == "maize", reg_plotter(YLDS, GDD)]  # P 0.04
regnew[CROP == "maize", reg_plotter(YLDS, KDD)]  # L 0.027
# srad
regnew[CROP == "maize", reg_plotter(YLDS, SSRT)]  # P 0.012 
regnew[CROP == "maize", reg_plotter(YLDS, SSRM)]  # Same, picked up wrong one
regnew[CROP == "maize", reg_plotter(YLDS, ASRT)]  # P 0.0288 
regnew[CROP == "maize", reg_plotter(YLDS, ASRM)]  # P 0.0299

# best predictors look to be
# MAP, SLOC (but polynomial relationship odd), SLHW, slcl, sbdm, all good
# STMX (linear KDD as a possible alternate), GDD, ASRM

#
regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4279

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4199

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(slhw, 2) + poly(sloc, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4044 # SLOC doesn't give as much value, NS on linear

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4293  # Doesn't add much

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4293  # Oct precip add a tiny bit, significant

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + JP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4312  # Jan precip adds more R2, but makes MAP linear NS

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + FP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4374  # Feb precip even more R2, but makes MAP linear NS

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
  summary(lm1)
  plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

# regnew[CROP == "maize", {
#   gam1 <- bam(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
#               poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + s(x, y), 
#              family = gaussian(link = "log"), method = "ML")
#   summary(gam1)
#   # plot(YLDS, predict(lm1))
# }]  # 0.4301  # Oct precip add a tiny bit, significant

@@

# regred[, c("Y10S", "Y90S") := as.integer(0)]

regmu <- regnew[CROP == "maize", lapply(.SD, mean), 
                by = list(CROP, WTH, Zone, SOL), 
                .SDcols = names(regnew)[c(3:4, 11:ncol(regnew))]]
regmu[, YST := "mu"]
# regnew[CROP == "maize" & Zone == "Zone2a" & WTH == "AABS" & 
#          SOL == "WI_CMZR003", mean(YLDS)]
qs <- function(x, y, ptile) {
  DF <- cbind(1:length(x), x, y)
  DF <- DF[order(DF[, 2]), ]
  DF[round(nrow(DF) * ptile), 3]
}

@@

a <- regnew[CROP == "maize" & Zone == "Zone3" & WTH == "AEQA" & 
            SOL == "WI_RGTZ054", YLDS]
b <- regnew[CROP == "maize" & Zone == "Zone3" & WTH == "AEQA" & 
            SOL == "WI_RGTZ054", mid]
qs(a, b, 0.9)
sort(a)

# Select out quantile values
m10 <- regred[,
  qs(YLDS, mid2, 0.1), by = list(CROP, WTH, Zone, SOL)
][, V1]
m90 <- regred[,
  qs(YLDS, mid2, 0.9), by = list(CROP, WTH, Zone, SOL) 
][, V1]

@@

regred2[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.071
regred2[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.039
regred2[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.020
regred2[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.017
regred2[CROP == "maize",  summary(lm(YLDS ~ poly(FP, 2)))]
regred2[CROP == "maize", reg_plotter(YLDS, 100 - slcl + slsi)]  # P 0.102
regred2[CROP == "maize", reg_plotter(YLDS, slcl)]  # P 0.1879
regred2[CROP == "maize", reg_plotter(YLDS, slsi)]  # P 0.03

regred2[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

regred2[CROP == "maize" & YST == "mu", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

regred2[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + poly(slcl, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

regred2[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(slhw, 2) + poly(sloc, 2) +
              poly(sbdm, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

regred2[CROP == "maize", {
  gam1 <- gam(YLDS ~ s(MAP) + s(JP) + s(slhw) + s(slcl) + (sbdm) + s(GDD) +
                KDD + s(ASRM), 
            family = gaussian(link = "log"), method = "ML")
}]

>>>>

p_root <- lmisc::set_base_path()
p_clim <- fp(p_root, "external/input_devel/climate")
p_yield <- fp(p_root, "external/input_devel/yield")
p_sgrids <- fp(p_root, "external/input_devel/soil")

@@

# Fit model
## Exploratory plots
```{r, eval = FALSE}
plot(prov)
regdt[CROP == "maize", points(x, y, pch = 16, cex = 0.5)]
# regdt[CROP == "maize", plot(MAP, YLDS)]
regdt[CROP == "maize" & YLDS > 9000, 
      points(x, y, pch = 16, cex = 0.5, col = "red")]
# regdt[CROP == "soybean", plot(MAP, YLDR)]
# regdt[CROP == "soybean" & YLDS > 1700, 
#       points(x, y, pch = 16, cex = 0.5, col = "blue")]

# Prune out yields > 8000 maize > 1700 soybean in subsistence case
# soyfilt <- regdt[CROP == "soybean" & YLDS > 1700, unique(.SD), 
#                 .SDcols = c("x", "y", "SOL", "WTH", "Zone")]
mzfilt <- regdt[CROP == "maize" & YLDS > 9000, unique(.SD), 
                .SDcols = c("x", "y", "SOL", "WTH")]
# regnew <- regdt[!x %in% soyfilt$x | !y %in% soyfilt$y]
regnew <- regdt[!x %in% mzfilt$x | !y %in% mzfilt$y]
regnew[CROP == "maize", points(x, y, pch = 16, cex = 0.5, col = "green")]

@@

# summary(lm(YLDS ~ poly(MAP, 2), data = regnew[CROP == "maize", ]))
# summary(lm(YLDR ~ poly(MSP, 2), data = regnew[CROP == "maize", ]))
regnew[CROP == "maize", plot(MAP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MAP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MAP, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MSP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "soybean", plot(MSP, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(FP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(MSP, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(MSP, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SLOC, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SLOC, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SLCL, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SLCL, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SLSI, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SLSI, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SBDM, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SBDM, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SLHW, YLDS, pch = 16, cex = 0.5)]  # pH looks pred
regnew[CROP == "maize", plot(SLHW, YLDR, pch = 16, cex = 0.5)]  # pH looks pred
regnew[CROP == "maize", plot(GDD, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(KDD, YLDS, pch = 16, cex = 0.5)]
# regnew[CROP == "maize", plot(log(KDD), YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(NKJ, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(NKF, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(ATMX, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(STMX, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(STMX, YLDR, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SSRT, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(SSRM, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(ASRT, YLDS, pch = 16, cex = 0.5)]
regnew[CROP == "maize", plot(ASRM, YLDS, pch = 16, cex = 0.5)]

@@

summary(lm(YLDS ~ poly(SBDM, 2) + poly(SLHW, 2), 
           data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(SLHW, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(SBDM, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(KDD, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDR ~ poly(STMX, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(STMX, 2), data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ KDD, data = regnew[CROP == "maize", ]))
summary(lm(YLDS ~ poly(GDD, 2), data = regnew[CROP == "maize", ]))

# p <- "SBDM"; d <- "YLDS"
# regnew[CROP == "maize", {
#   lm1 <- lm(get(d) ~ get(p))
#   lm2 <- lm(get(d) ~ poly(get(p), 2))
#   plot(get(p), get(d), pch = 16, cex = 0.1)
#   abline(lm1)
#   dum <- data.frame(floor(min(get(p))):ceiling(max(get(p))))
#   colnames(dum) <- p
#   lines(dum[, 1], predict(lm2, newdata = dum), col = "red")
#   summary(lm1)
#   # summary(lm2)
# }]

reg_plotter <- function(dep, pred) {
  lm1 <- lm(dep ~ pred)
  lm2 <- lm(dep ~ poly(pred, 2))
  plot(pred, dep, pch = 16, cex = 0.1)
  abline(lm1)
  prng <- c(floor(min(pred)), ceiling(max(pred)))
  dum <- data.frame(seq(prng[1], prng[2], by = diff(prng) / 100))
  # dum <- data.frame(floor(min(pred)):ceiling(max(pred)))
  colnames(dum) <- "pred"
  preds <- predict(lm2, newdata = dum)
  lines(dum[, 1], preds, col = "red")
  mtext(round(summary(lm1)$adj.r.squared, 4), side = 3, adj = 0)
  mtext(round(summary(lm2)$adj.r.squared, 4), side = 3, adj = 0.5)
}
# regnew[CROP == "maize", reg_plotter(YLDR, GDD)]  # 
regnew[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.058
regnew[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.050
regnew[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.0186
# regnew[CROP == "maize", reg_plotter(YLDS, JP - mean(JP))]  # P 0.012
regnew[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.009
regnew[CROP == "maize", reg_plotter(YLDS, OP)]  # P 0.007

regnew[CROP == "maize", reg_plotter(YLDS, SLHW)]  # P 0.16
regnew[CROP == "maize", reg_plotter(YLDS, SLOC)]  # P 0.43
# regnew[CROP == "maize", reg_plotter(YLDR, SLOC)]  # P 
regnew[CROP == "maize", reg_plotter(YLDS, SLCL)]  # P 0.191
regnew[CROP == "maize", reg_plotter(YLDS, SBDM)]  # 0.20
# mtax pairings
regnew[CROP == "maize", reg_plotter(YLDS, STMX)]  # P 0.07
regnew[CROP == "maize", reg_plotter(YLDS, ATMX)]  # P 0.044 
# NK
regnew[CROP == "maize", reg_plotter(YLDS, NKJ)]  # L 0.056
regnew[CROP == "maize", reg_plotter(YLDS, NKF)]  # P 0.024 
# GDD/KDD
regnew[CROP == "maize", reg_plotter(YLDS, GDD)]  # P 0.03
regnew[CROP == "maize", reg_plotter(YLDS, KDD)]  # L 0.026
# srad
regnew[CROP == "maize", reg_plotter(YLDS, SSRT)]  # P 0.015 
regnew[CROP == "maize", reg_plotter(YLDS, SSRM)]  # 0.0155
regnew[CROP == "maize", reg_plotter(YLDS, ASRT)]  # P 0.038 
regnew[CROP == "maize", reg_plotter(YLDS, ASRM)]  # P 0.039

# best predictors look to be
# MAP, SLOC (but polynomial relationship odd), SLHW, SLCL, SBDM, all good
# STMX (linear KDD as a possible alternate), GDD, ASRM

#
regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.2941

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.2926

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + poly(SLOC, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + poly(STMX, 2) + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.4244 # SLOC doesn't give as much value, NS on linear

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.2961  # Doesn't add much

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + JP + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.2972  # Jan precip adds more R2, but makes MAP linear NS

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + FP + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.303  # Feb precip even more R2, but makes MAP linear NS

regnew[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
  summary(lm1)
  plot(YLDS, predict(lm1))
}]  # 0.4301  # Oct precip add a tiny bit, significant

# regnew[CROP == "maize", {
#   gam1 <- bam(YLDS ~ poly(MAP, 2) + OP + poly(SLHW, 2) + poly(SLCL, 2) +
#               poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + s(x, y), 
#              family = gaussian(link = "log"), method = "ML")
#   summary(gam1)
#   # plot(YLDS, predict(lm1))
# }]  # 0.4301  # Oct precip add a tiny bit, significant

@@

# regred[, c("Y10S", "Y90S") := as.integer(0)]
regmu <- regred[CROP == "maize", lapply(.SD, mean), 
                by = list(CROP, x, y, WTH, SOL), 
                .SDcols = names(regred)[c(9:ncol(regred))]]
regmu[, YST := "mu"]
# regnew[CROP == "maize" & Zone == "Zone2a" & WTH == "AABS" & 
#          SOL == "WI_CMZR003", mean(YLDS)]
qs <- function(x, y, ptile) {
  DF <- cbind(1:length(x), x, y)
  DF <- DF[order(DF[, 2]), ]
  DF[round(nrow(DF) * ptile), 3]
}

@@

a <- regnew[CROP == "maize" & WTH == "AEQA" & SOL == "WI_RGTZ054", YLDS]
b <- regnew[CROP == "maize" & WTH == "AEQA" & SOL == "WI_RGTZ054", mid]
qs(a, b, 0.9)
sort(a)

# Select out quantile values
m10 <- regred[,
  qs(YLDS, mid2, 0.1), by = list(CROP, x, y, WTH, SOL)
][, V1]
m90 <- regred[,
  qs(YLDS, mid2, 0.9), by = list(CROP, x, y, WTH, SOL)
][, V1]

@@

regred2[CROP == "maize", reg_plotter(YLDS, MAP)]  # P 0.1235
regred2[CROP == "maize", reg_plotter(YLDS, MSP)]  # P 0.1097
regred2[CROP == "maize", reg_plotter(YLDS, JP)]  # P 0.036
regred2[CROP == "maize", reg_plotter(YLDS, FP)]  # P 0.026
regred2[CROP == "maize",  summary(lm(YLDS ~ poly(FP, 2)))]
regred2[CROP == "maize", reg_plotter(YLDS, 100 - SLCL + SLSI)]  # P 0.1243
regred2[CROP == "maize", reg_plotter(YLDS, SLCL)]  # P 0.19
regred2[CROP == "maize", reg_plotter(YLDS, SLSI)]  # P 0.049

regred2[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.3324  # Oct precip add a tiny bit, significant

regred2[CROP == "maize" & YST == "mu", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + OP + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2) + x + y)
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.3271  # Oct precip add a tiny bit, significant

regred2[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + poly(SLCL, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.34  # Oct precip add a tiny bit, significant

regred2[CROP == "maize", {
  lm1 <- lm(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + poly(SLOC, 2) +
              poly(SBDM, 2) + poly(GDD, 2) + KDD + poly(ASRM, 2))
  summary(lm1)
  # plot(YLDS, predict(lm1))
}]  # 0.466  # Oct precip add a tiny bit, significant

# Fit gams
gam1 <- gam(YLDS ~ s(MAP, k = 3) + s(JP, k = 3) + s(SLHW, k = 3) + 
              s(SLCL, k = 3) + s(SBDM, k = 3) + s(GDD, k = 3) + KDD + 
              s(ASRM, k = 3), family = gaussian(link = "log"), method = "ML", 
            data = regred2[CROP == "maize"])
gam2 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2), family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam3 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
              poly(ASRM, 2) + s(x, y), 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam4 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD + 
              s(x, y), family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam5 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(SBDM, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam6 <- gam(YLDS ~ poly(MAP, 2) + poly(JP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam7 <- gam(YLDS ~ poly(MAP, 2) + poly(SLHW, 2) + 
              poly(SLCL, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam8 <- gam(YLDS ~ poly(MAP, 2) + SLHW +  poly(SLCL, 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam9 <- gam(YLDS ~ poly(MAP, 2) + SLHW + poly(SLCL, 2) + 
              poly(100 - (SLCL + SLSI), 2) + poly(GDD, 2) + KDD, 
            family = gaussian(link = "log"), method = "REML", 
            data = regred2[CROP == "maize"])
gam10 <- gam(YLDS ~ poly(MAP, 2) + SLHW + poly(SLCL, 2) + 
               poly(SLOC, 2) + poly(GDD, 2) + KDD, 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam11 <- gam(YLDS ~ poly(MAP, 2) + SLHW + poly(SLCL, 2) + 
               SLOC + poly(GDD, 2) + KDD, 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam12 <- gam(YLDS ~ poly(MAP, 2) + SLHW + poly(SLCL, 2) + 
               SLOC + poly(GDD, 2) + KDD + s(x, y), 
             family = gaussian(link = "log"), method = "REML", 
             data = regred2[CROP == "maize"])
gam13 <- gam(YLDS ~ poly(MAP, 2) + poly(SLCL, 2) + 
